---
title: "Database Queries"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Database Queries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(unitcm)
```

This vignette demonstrates how to query each database module in the
UniTCM platform.

## Herb Explorer

### Search and filter herbs

```{r herbs-search}
# Text search
herbs <- search_herbs(q = "ginseng")

# Multi-value faceted filters
herbs <- search_herbs(
  flavors = c("sweet", "bitter"),
  properties = "warm",
  toxicity = "non-toxic"
)

# View available filter values
facets <- fetch_herb_facets()
facets$toxicity
facets$meridians
```

### Herb details and compounds

```{r herbs-detail}
herb <- get_herb("UNITCM_H001")
herb$herb_english_name
herb$efficacy

# Get compounds found in this herb
compounds <- get_herb_compounds("UNITCM_H001", all_pages = TRUE)
head(compounds)
```

### Export data

```{r herbs-export}
export_herbs(q = "ginseng", file = "ginseng_herbs.csv")
export_herb_compounds("UNITCM_H001", file = "ginseng_compounds.csv")
```

## Ingredient Explorer

### Search compounds by properties

```{r compounds-search}
# Drug-like compounds with molecular weight 200-500
compounds <- search_compounds(
  mw_min = 200, mw_max = 500,
  lipinski = "pass",
  is_drug = TRUE
)

# Get facet statistics
facets <- fetch_compound_facets()
facets$mw_range
```

### Compound details, ADMET, and targets

```{r compounds-detail}
# Full compound record
compound <- get_compound("UNITCM_I00001")
compound$component_name
compound$xref$pubchem_cid

# ADMET predictions (~90 endpoints)
admet <- get_compound_admet("UNITCM_I00001")
admet$caco2_permeability
admet$hia

# Predicted targets (DrugCLIP deep learning)
targets_dc <- get_compound_targets("UNITCM_I00001", method = "drugclip")

# ChEMBL similarity-based targets
targets_ch <- get_compound_targets("UNITCM_I00001", method = "chembl")

# Both sources combined
targets_all <- get_compound_targets("UNITCM_I00001", method = "both")
table(targets_all$source)
```

## Disease-Formula Atlas

### Search formulas by disease

```{r formulas-search}
# Search by text
formulas <- search_formulas(q = "insomnia")

# Filter by ICD-11 classification
formulas <- search_formulas(
  level1 = "Neoplasms",
  mapping_confidence = c("high", "medium")
)

# Browse the disease classification tree
tree <- fetch_disease_tree()

# Available filter options
list_book_sources()
list_origin_sources()
list_dosage_forms()
```

### Formula details and doses

```{r formulas-detail}
formula <- get_formula(1)
formula$formula_name
formula$efficacy

# Herb composition and dosage
doses <- get_formula_doses(1)
doses
```

## TCM Ontology

### Search and navigate the ontology

```{r ontology-search}
# Full-text search
results <- search_ontology("Qi stagnation")

# Get entity with relations
entity <- get_ontology_entity("TCM_0001")
entity$ancestors
entity$children
entity$external_mappings

# Navigate the hierarchy
children <- get_ontology_children("TCM_0001")
ancestors <- get_ontology_ancestors("TCM_0001")
descendants <- get_ontology_descendants("TCM_0001", max_level = 2)
```

### Ontology tree and statistics

```{r ontology-tree}
# Fetch tree at depth 2
tree <- fetch_ontology_tree(depth = 2)

# Statistics
stats <- fetch_ontology_stats()
stats$total_entities

# Top-level categories
categories <- list_ontology_categories()
```

### Cross-database mapping

```{r ontology-mapping}
# Find TCM entities mapped to a MeSH term
mapped <- search_ontology_mapping("MeSH", "D008516")

# Export ontology
export_ontology("csv", file = "tcm_ontology.csv")
export_ontology("json", depth = 3, file = "tcm_ontology.json")
```

## MIDAS Gene-Disease Analysis

### Gene-to-disease and disease-to-gene

```{r midas-basic}
# Which diseases are associated with these genes?
g2d <- query_gene_diseases(c("TP53", "BRCA1", "EGFR"))
g2d
attr(g2d, "gene_mapping")

# Which genes are associated with breast cancer?
d2g <- query_disease_genes("breast cancer")
d2g
attr(d2g, "matched_diseases")
```

### Disease enrichment analysis

```{r midas-enrichment}
gene_list <- c("TP53", "BRCA1", "EGFR", "VEGFA", "MYC", "KRAS")
enrichment <- query_disease_enrichment(
  gene_list,
  p_value_cutoff = 0.05,
  correction_method = "fdr"
)
enrichment
attr(enrichment, "total_significant")
```

### Utility functions

```{r midas-utils}
# Convert mixed gene identifiers
convert_gene_ids(c("TP53", "7157", "ENSG00000141510"))

# Autocomplete disease names
autocomplete_disease("breast")

# Available data sources
fetch_midas_sources()

# Database statistics
fetch_midas_stats()
```
